High-order cognition is supported by complex but compressible brain activity patterns

### 2023-02-20 paper/main.tex
draft of abstract

We applied dimensionality reduction algorithms to the activity
patterns in each experimental condition. Specifically, we sought to understand
the “dimensionality” of the neural patterns that were sufficient to decode
participants’ listening times (or approach was similar to that of Mack et al.
2017). We found that even low-dimensional embeddings of the data were
sufficient to accurately decode listening times from the intact story
recording, whereas finer temporal scramblings of the story required
higher-dimensional embeddings of the data to reach peak decoding accuracy.

>>>>

We applied dimensionality reduction algorithms and pattern classifiers to
functional neuroimaging data collected as participants listened to a story,
temporally scrambled versions of the story, or underwent a resting state
scanning session. These experimental conditions were intended to require
different depths of processing and insprire different levels of engagement. We
considered two primary aspects of the data. First, we treated the number of
features (components) required to achieve a threshold decoding accuracy as a
proxy for the ``compressibility'' of the neural patterns (where fewer
components indicate higher compressibility). Second, we treated the maximum
achievable decoding accuracy across participants as an indicator of the
``stability'' of the recorded patterns. Overall, we found that neural patterns
recorded as participants listened to the intact story required fewer features
to achieve comparable classification accuracy to the other experimental
conditions. However, the peak decoding accuracy (achievable with more features)
was also highest during intact story listening. Taken together, our work
suggests that our brain networks flexibly reconfigure according to ongoing task
demands, and that the activity patterns associated with higher-order cognition
and high engagement are both more complex and more compressible than the
activity patterns associated with lower-order tasks and lower levels of
engagement.
